R and Python clients for GBIF species occurrence data

### 2017-08-10 manuscript.tex
minor tweaks to ms, added many references

Dataset metrics are another useful way of figuring out what datasets you
may want to use. One drawback is that these metrics data are only
available for datasets of type \emph{checklist}, but there are quite a
lot of them (16057).

@@

The GBIF backbone taxonomy is used in GBIF to have a consistent way to
refer to taxonomic names throughout their services. The backbone has
5194833 unique names and 2420842 species names. The backbone taxonomy is
also a dataset with key \texttt{d7dddbf4-2cf0-4f39-9b2a-bb099caae36c}
(\url{http://www.gbif.org/dataset/d7dddbf4-2cf0-4f39-9b2a-bb099caae36c}).

@@

The \texttt{rgbif} and \texttt{pygbif} libraries provide programmatic
interfaces to GBIF's application programming interface (API) - a
powerful tool for working with species occurrence data, and facilitating
reproducible research. In fact, the \texttt{rgbif} package has already
been used in 22 scholarly publications (as of 2015-11-14).

@@

@@

@@

@@

@@

@@

@@

>>>>

Dataset metrics are another useful way of figuring out what datasets you
may want to use. One drawback is that these metrics data are only
available for datasets of type \emph{checklist}, but there are quite a
lot of them (16084).

@@

The GBIF backbone taxonomy is used in GBIF to have a consistent way to
refer to taxonomic names throughout their services. The backbone has
5296599 unique names and 2525274 species names. The backbone taxonomy is
also a dataset with key \texttt{d7dddbf4-2cf0-4f39-9b2a-bb099caae36c}
(\url{http://www.gbif.org/dataset/d7dddbf4-2cf0-4f39-9b2a-bb099caae36c}).

@@

The \texttt{rgbif} and \texttt{pygbif} libraries provide programmatic
interfaces to GBIF's application programming interface (API) - a
powerful tool for working with species occurrence data, and facilitating
reproducible research. In fact, the \texttt{rgbif} package has already
been used in 22 scholarly publications (as of 2016-08-10) (Amano,
Lamming \& Sutherland, 2016, Bartomeus et al. (2013), Barve (2014), Bone
et al. (2015), Collins et al. (2015), Drozd \& {Š}ipo{š} (2013), Kong,
Huang \& Duan (2015), Richardson, Roux \& Wilson (2015), Turner,
Fr{é}ville \& Rieseberg (2015), Verheijen et al. (2015), Zizka \&
Antonelli (2015), Butterfield et al. (2016), Dellinger et al. (2015),
Feitosa et al. (2015), Malhado et al. (2015), Werner et al. (2015),
Robertson, Visser \& Hui (2016), Davison et al. (2015), Janssens et al.
(2016)).

@@

Amano T., Lamming JDL., Sutherland WJ. 2016. Spatial gaps in global
biodiversity information and the role of citizen science.
\emph{BioScience} 66:393--400.

Bartomeus I., Park MG., Gibbs J., Danforth BN., Lakso AN., Winfree R.
2013. Biodiversity ensures plant-pollinator phenological synchrony
against climate change. \emph{Ecology Letters} 16:1331--1338.

Barve V. 2014. Discovering and developing primary biodiversity data from
social networking sites: A novel approach. \emph{Ecological Informatics}
24:194--199.

@@

Bone RE., Smith JAC., Arrigo N., Buerki S. 2015. A macro-ecological
perspective on crassulacean acid metabolism (CAM) photosynthesis
evolution in afro-madagascan drylands: Eulophiinae orchids as a case
study. \emph{New Phytologist} 208:469--481.

@@

Butterfield BJ., Copeland SM., Munson SM., Roybal CM., Wood TE. 2016.
Prestoration: Using species in restoration that will persist now and
into the future. \emph{Restor Ecol}.

@@

Collins R., Ribeiro ED., Machado VN., Hrbek T., Farias I. 2015. A
preliminary inventory of the catfishes of the lower rio nhamundá, brazil
(ostariophysi, siluriformes). \emph{BDJ} 3:e4162.

Davison J., Moora M., Opik M., Adholeya A., Ainsaar L., Ba A., Burla S.,
Diedhiou AG., Hiiesalu I., Jairus T., Johnson NC., Kane A., Koorem K.,
Kochar M., Ndiaye C., Partel M., Reier U., Saks U., Singh R., Vasar M.,
Zobel M. 2015. Global assessment of arbuscular mycorrhizal fungus
diversity reveals very low endemism. \emph{Science} 349:970--973.

Dellinger AS., Essl F., Hojsgaard D., Kirchheimer B., Klatt S., Dawson
W., Pergl J., Py{š}ek P., Kleunen M van., Weber E., Winter M., Hörandl
E., Dullinger S. 2015. Niche dynamics of alien species do not differ
among sexual and apomictic flowering plants. \emph{New Phytologist}
209:1313--1323.

Drozd P., {Š}ipo{š} J. 2013. R for all (i): Introduction to the new age
of biological analyses. \emph{Casopis slezskeho zemskeho muzea (A)} 62.

@@

Feitosa YO., Absy ML., Latrubesse EM., Stevaux JC. 2015. Late quaternary
vegetation dynamics from central parts of the madeira river in brazil.
\emph{Acta Bot. Bras.} 29:120--128.

@@

Janssens SB., Vandelook F., Langhe ED., Verstraete B., Smets E.,
Vandenhouwe I., Swennen R. 2016. Evolutionary dynamics and biogeography
of musaceae reveal a correlation between the diversification of the
banana family and the geological and climatic history of southeast asia.
\emph{New Phytologist} 210:1453--1465.

Kong X., Huang M., Duan R. 2015. SDMdata: A web-based software tool for
collecting species occurrence records. \emph{PLOS ONE} 10:e0128295.

Malhado AC., Oliveira-Neto JA., Stropp J., Strona G., Dias LC., Pinto
LB., Ladle RJ. 2015. Climatological correlates of seed size in amazonian
forest trees. \emph{J Veg Sci} 26:956--963.

@@

Richardson DM., Roux JJL., Wilson JR. 2015. Australian acacias as
invasive species: Lessons to be learnt from regions with long planting
histories. \emph{Southern Forests: a Journal of Forest Science}
77:31--39.

Robertson MP., Visser V., Hui C. 2016. Biogeo: An r package for
assessing and improving data quality of occurrence record datasets.
\emph{Ecography} 39:394--401.

Turner KG., Fr{é}ville H., Rieseberg LH. 2015. Adaptive plasticity and
niche expansion in an invasive thistle. \emph{Ecol Evol} 5:3183--3197.

Verheijen LM., Aerts R., Bönisch G., Kattge J., Bodegom PMV. 2015.
Variation in trait trade-offs allows differentiation among predefined
plant functional types: Implications for predictive ecology. \emph{New
Phytologist} 209:563--575.

Werner GDA., Cornwell WK., Cornelissen JHC., Kiers ET. 2015.
Evolutionary signals of symbiotic persistence in the legumerhizobia
mutualism. \emph{Proceedings of the National Academy of Sciences}
112:10262--10269.

Zizka A., Antonelli A. 2015. \emph{speciesgeocodeR: An r package for
linking species occurrences, user-defined regions and phylogenetic trees
for biogeography, ecology and evolution}. Cold Spring Harbor Laboratory
Press.
